### 2018-03-15 paper.Rmd
added architecture section

Database Architecture
MySQL
Extensibility
Versioning

## Cross-linguistic organization

>>>>

Unlike the *parsing* approach in CLAN which entails the sequential processing of strings, childes-db treats CHILDES as a set of linked tables, with records corresponding to intuitive abstractions such as words, utterances, and transcripts. Users of data analysis languages like R or Julia, libraries like Pandas, or those familiar with Structured Query Language (SQL) will be familiar with operations on tables such as filtering (subsetting), sorting, aggregation (grouping), and joins (merges). We expect most users to interact with the database through the childesr API for the R language described below, we begin with a top-level description and motivation for the design of the database schema, details of the database’s current technical implementation, and our versioning scheme.

At its core, childes-db is a database consisting of a set of linked tabular data stores where rows corresponding to records. The smallest unit of abstraction tracked by the database is a token, treated here as the citation orthographic form of a word. Using the standardized written form of the word facilitates the computation of lexical frequency statistics, for comparison or aggregation across children or time periods. Deviations from the citation form—reflecting the deviations from citation forms that are particularly common in the course of language development— are kept as a separate (possibly null) field associated with each token.

Other tables in the database comprise a hierarchical system of collections built on top of tokens—*utterance*, *transcription*, *corpus*, and *collection*—which stores attributes appropriate for each level of description. Every entity includes attributes that link it to all higher-order collections, e.g., an utterance lists the transcript, corpus, and collection to which it belongs.  An *utterance* contains one or more words and includes fields such as the utterance type (declarative, question, etc.), total number of tokens, and the total number one morphemes. [[ How are we handling the long tail of tiers?]]. A *transcription* consists of one or more utterances and includes the date collected, the name of the target child and age in days (if defined), and the filename from CHILDES. A *corpus* consists of one or more transcripts, corresponding to well-known collections like the Brown or Providence corpus. Finally, a *collection* is a superordinate collection of corpora generally corresponding to a geographic region, following the convention in CHILDES. Because every entity can be linked to a top-level collection, childes-db places content from all languages in the same table for a given level of abstraction.

We chose to treat participants separate from the above hierarchy because children can occur across transcripts (e.g., the Brown Corpus). A participant is associated with every word and utterance, including a name, role, 3-letter CHILDES identifier (CHI, MOT, FAT, etc.), and the range of ages (or age of corresponding child) for which they are observed. For non-child participants, the record contains an identifier for the target child to associate caregivers and children.
Technical Implementation

The above tabular datastore is implemented as a MySQL database.  As an industry-standard, open-source relational database server, the above database can be accessed directly from a wide range of programming languages. The childes-db project provides hosted read-only databases for direct access and for `childesr` (described below) as well as compressed .sql exports for local installation. While the former is appropriate for casual users, local installation can provide performance gains (allowing a user to access the database on their local machine or on their network) as well as write support in case users wish to store derived information in the same database.
